---
title: "p8105_hw1_ao2965"
author : Alissa Shams Orchi
output: github_document
date: "2025-09-19"
---

```{r}
library(moderndive)
data("early_january_weather")
```
The variables are as follows:

* origin: Weather station. Named origin to facilitate merging with nycflights13::flights data.

* year, month, day, hour: Time of recording.

* temp, dewp: Temperature and dewpoint in F.

* humid: Relative humidity.

* wind_dir, wind_speed, wind_gust: Wind direction (in degrees), speed and gust speed (in mph).

* precip: Precipitation, in inches.

* pressure: Sea level pressure in millibars.

* visib: Visibility in miles.

* time_hour: Date and hour of the recording as a POSIXct date.

* The mean temperature is `r mean(early_january_weather$temp)` °F. 

* This dataset has `r nrow(early_january_weather)` rows.

* This dataset has `r ncol(early_january_weather)` columns.


```{r}
library(ggplot2)
ggplot(early_january_weather, aes(x = time_hour, y = temp,color = humid)) + geom_point() + labs(
    title = "Temperature over Time (colored by Humidity)",
    x = "time_hour", y = "temp", color = "Humidity"
  )

```


The pattern of the scatterplot shows that as time goes by, temperature is seen to be increasing. The plot also shows that humidity is recorded to be increasing with time. There is a period of time around Jan 14 when humidity is recorded to be higher than usual. 
```{r}
ggsave("scatter_plot.pdf", height = 4, width = 6)
```


2.

```{r message= FALSE}
library(tidyverse)
set.seed(1234)
q2_df = tibble(
  q2_num = rnorm(10),                                    
  q2_log = q2_num > 0,                                   
  q2_chr = c("apple","banana","cherry","grape","blueberry","blackberry", "strawberry", "rasberry", "mango", "kiwi"), 
  q2_vec = factor(rep(c("high", "medium", "low"), length.out = 10))      
)
q2_df
```

```{r}
mean(pull(q2_df, q2_num))

```

The mean for the random sample of size 10 from a standard Normal distribution was -0.3831574

```{r}
mean(pull(q2_df, q2_log))

```

The mean for logical vector indicating whether elements of the sample are greater than 0 was 0.4
```{r}
mean(pull(q2_df, q2_chr))
mean(pull(q2_df, q2_vec))
```

We could not find the mean for a character vector of length 10 and a factor vector of 
length 10, with 3 different factor “levels”. This is because means only work for numerical values.
```{r results= 'hide', warning=FALSE}
as.numeric(q2_df %>% pull(q2_log))  
as.numeric(q2_df %>% pull(q2_chr))  
as.numeric(q2_df %>% pull(q2_vec))

```
Logical variables, which contain TRUE and FALSE values, are internally stored as 1 and 0. So, converting logicals to numeric results in 1s and 0s, making it possible to calculate the mean as the proportion of TRUE values. Character variables consist of strings that do not represent numbers. Applying as.numeric() to these produces NA values with a warning because characters like fruit names cannot be meaningfully converted to numbers. Factor variables are categorical but internally stored as integer codes representing the factor levels. When converted to numeric, the codes will give mean, but it will give mean for the integer level codes. So, it would be meaningless.

This explains why the mean function in R works directly on numeric and logical types but produces errors when applied to character or factor variables. It is also important to be careful of averaging mixed columns, or we will end up with misleading numbers.
